---
title: "Week 4 Supplement"
author: "BL Mosby"
date: "2/6/2021"
output: html_document
---

```{r} 
r = getOption("repos")
r["CRAN"] = "http://cran.us.r-project.org"
options(repos = r)

library(tidyverse)
library(haven)
library(readxl)

```


## Correction to the Excel download from Async
Here are a few tweaks to the async code for the Excel file download that may run a bit smoother:
```{r}
## Web page: 
##http://nces.ed.gov/programs/digest/d14/tables/dt14_204.10.asp

#If you have been getting an error with the download, add the mode argument (which directs RStudio to download as a binary file).
if(file.exists("free.xls")==FALSE){
  download.file("http://nces.ed.gov/programs/digest/d14/tables/xls/tabn204.10.xls",destfile="free.xls", mode="wb")
free<-read_excel("free.xls",skip=4,col_names=FALSE)  
}else{
  free<-read_excel("free.xls",skip=4,col_names=FALSE)
}

head(free)
```

Next, to get rid of the empty rows and columns, you may need to adjust the column names to match what's in your dataframe.
```{r}
# Now need to clean up 
#Get rid of unwanted columns

free2<-free[ ,-(c(3,6,9,12,15,18))]

#Get rid of unwanted rows-- My column titles are "...1" so I will change the is.na argument to reflect that:
free2<-free2%>%filter(is.na(...1)==FALSE)
```
The rest of the code chunks for the Excel download should work as is.


---

## Zipped Files
Some data files that are very large are compressed into a zipped file (along with codebooks or other metadata files). These take a two step process, where you must also know the name of the data file within the zipped folder.

```{r}
#creating a temporary file to hold the zipped folder
temp <- tempfile()

#downloading the zipped folder from the data source webpage, and unzipping the SPSS data file
download.file("https://gss.norc.org/Documents/spss/GSS_2014_Merged_SPSS.zip", temp)
unzip(temp, "gss2014merged_r10 - spss.sav")

#reading in the spss command and closing the connection
gss14<-read_spss("gss2014merged_r10 - spss.sav")
unlink(temp)

head(gss14)

```


## Homework 4 Exercises
For Assignment 4, replace #2 and #3 with the following:

2. U.S. National Election Survey data from 2004. Save as file name `nes.RData`. [https://osf.io/zpuhd/download]("https://osf.io/zpuhd/download")

3. General Social Survey (GSS), pick any of the individual year datasets (1972 - 2018). Save as file name `gss<year>.Rdata`. Note this will be an SPSS file within a zipped file, and you will need to use the code above.
[https://gss.norc.org/get-the-data/spss](https://gss.norc.org/get-the-data/spss).


## Extra Excel data
My code for the additional Excel dataset from async. (Note, there is more than one correct way to go about cleaning the data.)
```{r}
## Web page: 
##http://nces.ed.gov/programs/digest/d14/tables/dt14_204.10.asp

download.file("http://nces.ed.gov/programs/digest/d14/tables/xls/tabn302.10.xls",destfile="newxl.xls", mode="wb")

newxl<-read_excel("newxl.xls",skip=6,col_names=FALSE)

#Get ride of empty rows and unneeded columns.
newxl<-newxl[1:54,]
newxl<-newxl[ ,-(c(3,5,7,9,11,13,15,17,19,21,23,25))]

#Add useful names to the columns.
names(newxl)<-c("year", "hscomptotal", "hscompmales","hscompfemales","incollegetotal","total2y","total4y",  "collegetotalmales", "males2y",  "males4y", "collegetotalfemales",  "females2y",  "females4y")

#Change the year variable into a format that R can understand and drop the old one.
yeardate = seq(from = as.Date("1960-01-01"), to = as.Date("2013-01-01"), by = 'year')

newxl2<-cbind(newxl,yeardate)
newxl2<-newxl2[,-1]

head(newxl2)
```

This was just a graph I did for fun, to visualize one of the variables.
```{r}
gg<-ggplot(newxl2, aes(x=yeardate, y=incollegetotal))
gg<-gg+geom_point()
gg
```

