---
title: "Plots for Classifiers"
output: html_document
---

Plots are rarely used in the context of classification, but they can aid understanding the behavior of the dependent variable. I'll show three ways of thinking about plots for classification: bar graphs, heatmaps, and plotting the probability predictions from a logit model.


We'll continue working with the Random Acts of Pizza dataset. 

```{r}
library(tidyverse)
library(modelr)
library(knitr)
load("za.RData")
```

We always want to start with a cross tab of our dependent variable as a function of other variables. We structure crosstabs in a very particular way for the purposes of research: the independent variable goes on the rows, and the dependent variable goes on the columns. If proportions or percentages are going to be calculated, they should be calculated across rows. 

Recalling our previous lesson, let's look at a crosstab of the dependent, binary variable `got_pizza` with the independent variable of `student`

```{r}
#Note the order: rows first (independent variable), followed by columns (dependent variable).
tab_student<-with(za,table(student,got_pizza))
tab_student
```

If we want to make this a little better, we can change the row and column titles

```{r}
colnames(tab_student)<-c("No Pizza","Received a Pizza")
tab_student

kable(tab_student) #recall that a kable is a markdown formatted table, and particularly recommended for use in reports
```

If we want to add proportions to this table, we can do that like so:
```{r}
tab_student_prop<-prop.table(tab_student,margin=1)
kable(tab_student_prop)
```

Sometimes (okay, all the times) audiences prefer percentages. Easy enough to do:
```{r}
kable(round(tab_student_prop*100,2))
```

If you want to include a third variable in a crosstab, that requires splitting the dataset. For instance, if we want to know the proportion of posts that include "student" AND "grateful" that received pizza, we would do this:

```{r}
#filter only data that had the word grateful in the post:
tab_student_grateful<-with(filter(za,
                           as.character(grateful)=="Grateful in post"),
                           table(student,got_pizza))
tab_student_grateful
```
Note that this table represents *only* the posts that contained "grateful."

*Outcome by "Student" AND "Grateful"*
```{r}
prop.table(tab_student_grateful,margin=1)%>%kable()
```


## Barcharts

To format the data for barcharts, we make use of the concept of conditional means. Let's use two independent variables to calculate the probability of receiving a pizza: poor and student. 

```{r}
#This is similar to code from weeks 3 and 4 for calculating and graphing conditional means
za_sum<-za%>%group_by(poor,student)%>%dplyr::summarize(prob_pizza=mean(got_pizza,na.rm=TRUE))

gg1<-ggplot(za_sum,aes(x=student,y=prob_pizza,fill=poor))
gg1<-gg1+geom_bar(stat="identity",position="dodge")
gg1<-gg1+xlab("Student Status")+ylab("Pr(Pizza)")
gg1<-gg1+theme(legend.title=element_blank())
gg1<-gg1+geom_text(aes(label=round(prob_pizza,2)),
                   position=position_dodge(width=.9),
                   vjust=-.25)
gg1
```
Which group was most likely to get pizza? Least likely?

## Heat Maps

```{r}
za<-za%>%mutate(score_decile=ntile(score,5),
                karma_decile=ntile(karma,5))


za_sum<-za%>%group_by(score_decile,karma_decile)%>%
  dplyr::summarize(prob_pizza=mean(got_pizza,na.rm=TRUE))%>%
  arrange(-prob_pizza)

za_sum<-za_sum%>%dplyr::filter(!(is.na(score_decile)),!(is.na(karma_decile)))

#The usual setup for a gg plot that specifies the data, x- and y-values, and the fill aesthetics
gg<-ggplot(za_sum,
           aes(x=as.factor(score_decile),
               y=as.factor(karma_decile),fill=prob_pizza))
gg<-gg+geom_tile() #This geometry will plot the date as a grid, where the color of each cell corresponds to the likelihood of getting pizza
gg<-gg+scale_fill_gradient(low="white",high="red")
gg<-gg+xlab("Emotional Valence")+ylab("Reddit Karma")
gg<-gg+theme(legend.title=element_blank())
gg

```


## Plotting by probabilities from models
```{r}
#Definining a logistic regression model with independent variables karma, total posts, RAOP posts, student, and grateful
logit_mod<-glm(got_pizza~
             karma+
             total_posts+
             raop_posts+
             student+
             grateful,
             data=za,
            na.action=na.exclude,
            family=binomial(link="logit"),
               y=TRUE)

```



```{r}
#Creating a hypothetical, representative dataset to create a graphical representation of the model
hypo_data<-data_grid(za, 
                     total_posts=seq_range(total_posts,n=100),
                                           karma=mean(karma,na.rm=TRUE),
                                           raop_posts=mean(raop_posts,na.rm=TRUE),
                     grateful=levels(grateful)[1],
                                           student=levels(student))%>%
              mutate(pred=predict(logit_mod,newdata=.,type="response"))

gg<-ggplot(hypo_data,
           aes(x=total_posts,y=pred,color=student))
gg<-gg+geom_line()
gg<-gg+xlab("Total Posts")+ylab("Predicted Probability of Pizza")
gg
```

